---
title: "Analysis of BiPOCD data"
author: "Sebastian Sauer"
date: "4 Juli 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(stringr)
library(ggvis)
library(pander)
library(purrr)
library(caret)
library(broom)
library(ggstance)
library(knitr)
library(gridExtra)
library(tibble)



# define paths
path_data <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/raw_data"
path_file_data <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/raw_data/BiPOCD prediction.csv"
path_code <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/R_code"
path_figs <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/Figures"
path_obj <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/data_objects"

# load data
data <- read_csv(path_file_data)


# constants
red1 <- "#880011"
```




```{r funs}

gg_bar <- function(x, ...){
  { 
    ggplot(data_frame(x), aes(x = x)) + 
      geom_bar() +
      coord_flip() 
  }
}


add_na_col <- function(x){
   mutate(x, na = 0)
}

has_n_col <- function(x, n = 6){
  return(ncol(x) == n)
}

```



# First glimpse

```{r}
glimpse(data)
```


##  Variable names
The main (raw) data file has a dimension of `r dim(data)` (rows*cols). The col names are:

```{r}
names(data)
```

# Overviews
## Summaries of quantitative variables

```{r}


data %>% 
  select_if(is.numeric)  %>% 
  lapply(., function(x) tidy(summary(x))) %>%   # compute tidy summary of each var
  map_if(., has_n_col, add_na_col) %>%   # add na-col if missing
  do.call(rbind, .) -> sum_num  # bind list elements into df

kable(sum_num)

```


## Histograms of quantitative variables

```{r histograms_numeric, cache = TRUE, fig.width = 16, figh.height = 20}
data %>% 
  select_if(is.numeric) %>% 
  gather(key = variable, value = value) %>% 
  ggplot(aes(x = value)) + 
  geom_histogram() + 
  facet_wrap(~variable, scales = "free", ncol = 5)

```


## Histogram of non-numeric variables

```{r, cache = TRUE}

data %>% 
  select_if(negate(is.numeric)) %>% 
  lapply(., function(x) gg_bar(x)) -> gg_bar_list

do.call(grid.arrange, gg_bar_list)

```


## Check for (Near) Zero variance (nz/nzv) variables

Predictors without (much) variance do not add to the model. Better exclude them, at least know them.

```{r}
data %>% 
  select(-ID) %>% 
  select_if(is.numeric) %>% 
  nearZeroVar(., saveMetrics = TRUE) %>% 
  pander

data %>% 
  select(-ID) %>% 
  select_if(is.numeric) %>% 
  nearZeroVar(., saveMetrics = TRUE) %>% 
  filter(nzv == TRUE) %>% 
  pander
```


## Check for highly correlated variables
Highly correlated (>.90) do not provide more information, better exclude them.

```{r}

data %>% 
  select_if(is.numeric) %>% 
  cor -> data_cor

highCorr <- sum(abs(data_cor[upper.tri(data_cor)]) > .90)
highCorr
```

No variables are highly correlated (>.90) in this dataset.


## Missing values

```{r}
 sum_num %>% 
  select(na) %>% 
  rownames_to_column %>% 
  filter(na != 0) %>% 
  ggplot(aes(x  = rowname, y = na, label = na)) + geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Number of missing values per variable")
```

