---
title: "BiPOCD_modeling"
author: "Sebastian Sauer"
date: "11 Juli 2016"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: true
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE)


library(knitr)
library(readr)
library(tidyr)
library(caret)
library(ggplot2)
library(tibble)
library(dplyr)



# define paths
path_data <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/raw_data"
path_file_data <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/raw_data/BiPOCD_raw_data.csv"
path_code <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/R_code"
path_figs <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/figs"
path_obj <- "~/Documents/OneDrive/Forschung/Online_Behavior_Therapy/data_objects"

# load data
data <- read_csv(path_file_data)


# constants
red1 <- "#880011"


# some important sub data sets

search_string <- "sum"
data %>% 
  select(contains("sum")) -> data_sumscores


data %>% 
  mutate(responder_3m_f = factor(responder_3m)) -> data

data$responder_3m_f <- relevel(data$responder_3m_f, ref = "1")


data_sumscores$responder_3m_f <- data$responder_3m_f
data_sumscores$CYBOCS_3m <- data$CYBOCS_3m



data %>%
  nearZeroVar(., saveMetrics = TRUE) %>%
  rownames_to_column() %>%
  filter(nzv == TRUE) -> data_nzv

nzv_vars <- data_nzv$rowname


data %>%
  select(Depression:GAD, responder_3m, responder_3m_f) %>%
  select(-one_of(nzv_vars)) %>%
  na.omit -> data_comorb




# data_short -- variables used for modeling


data %>%
  caret::nearZeroVar(., saveMetrics = TRUE) %>%
  tibble::rownames_to_column() %>%
  filter(nzv == TRUE) -> data_nzv

nzv_vars <- data_nzv$rowname

data %>% 
  select(dplyr::contains("sum"),  # sumscores of scales
         Depression:numberdiagnos,  # comorbidity
         sex:OCD_treatm_exp,  # demographics
         responder_3m_f, CYBOCS_3m,  # outcome
         -one_of(nzv_vars)) ->  # exclude nzv variables
  data_mod


# write_csv(data_mod, path = "data_mod.csv")



# list with modeling results

mod_results <- vector(mode = "list")


```



# Replace/omit missing values


How many missings do we have?

```{r}
data_mod %>% select_if(function(col) sum(is.na(col)) != 0) %>% names


data_mod %>% 
  select_if(function(col) sum(is.na(col)) != 0) %>% 
  summarise_each(funs(sum(is.na(.)))) %>% 
  kable
```

Oh no! Some NA's in our outcome variable, that's is a sad story. That reduces the sample size. Ok, then be hard now. At least the 5 missings occur in both variables for the same cases. Good bye, guys.

```{r}
data_mod %>% 
  filter(!is.na(responder_3m_f), !is.na(CYBOCS_3m)) -> data_mod
```

The dimension of our data set is now `r dim(data_mod)`.


What about the rest of the variables with NAs? They do have quite a number of NAs, and the variables do not appear so important. So let's omit these variables.

```{r}
data_mod %>% 
  dplyr::select(-OCDonset, -yearswithOCD) -> data_mod
```

Now the dimension of our data set is: `r dim(data_mod)`.


# Separate train and test dataset

Let's split the data set in two parts, for training (80%) and for testing, to avoid bias from overfitting.

```{r}
set.seed(42)
trainIndex <- createDataPartition(data_mod$responder_3m_f, p = .7,
                                  list = FALSE,
                                  times = 1)

train <- data_mod[trainIndex, ]
test <- data_mod[-trainIndex, ]
```



# Factor variables with sparsely populated levels -- exclude

Some variables are sparsely populated, e.g.:

```{r}
 data_mod %>% 
  select(medication) %>% 
  group_by(medication) %>% 
  summarise(n = n()) %>%
  kable
```


Some probems occured with the sparse levels. Better exclude it.

```{r}

sparse_vars <- c("medication", "Education_parent", "OCD_treatm_exp", "Birthcountry", "treatm_exp")

data_mod %>% 
  dplyr::select(-one_of(sparse_vars)) -> data_mod


```

The factor variables with sparse levels (now excluded) are: `r sparse_vars`.




# classification (dichotomous)

As we have two outcomes, one dichotomous and one metric, we can both do classification and regression. Let's look at classification first.


Before I forget: we need to exclude the metric outcome variable.

```{r}
data_mod %>% 
  select(-CYBOCS_3m) -> data_class

set.seed(42)
trainIndex <- createDataPartition(data_class$responder_3m_f, p = .7,
                                  list = FALSE,
                                  times = 1)

train <- data_class[trainIndex, ]
test <- data_class[-trainIndex, ]


```


## LM

Let's fit a linear logistic model. The model showed some problems with factor variables that have sparsely populated levels. 




```{r}

ctrl <- trainControl(method = "repeatedcv", repeats = 3)

glmfit1 <- train(responder_3m_f ~ .,
                 data = train,
                 method = "glm",
                 trControl = ctrl,
                 preProc = c("center", "scale"))


glmfit1_pred <- predict(glmfit1, newdata = test)

(mod_results[[1]] <- confusionMatrix(data = glmfit1_pred, test$responder_3m_f))
mod_results[[1]]$name <- "glmfit1"
mod_results[[1]]$varimp <- varImp(glmfit1, scale = FALSE)

```

Classification rate was ...bad...? But now look at the variable importance. Note: 
 >Linear Models: the absolute value of the t–statistic for each model parameter is used. 

(From caret help `varimp`.)


```{r lm1_varimp}
mod_results[[1]]$varimp[[1]] %>% 
  select(everything()) %>% 
  rownames_to_column %>% 
  arrange(desc(Overall))%>% 
  top_n(5) %>% 
  kable
```

Only top-5 shown.

